Poincar\'e--Friedrichs constants over domains with shellable triangulations

### 2024-06-26 main.tex
Update on Overleaf.

\begin{abstract}
    We construct upper bounds for Poincar\'e--Friedrichs constants over triangulated domains. 
    These includes lower estimates for the singular values of the gradient, curl, and divergence operators as special cases. 
    The main application that we envision are computable estimates over finite element patches. 
\end{abstract}

@@

\paragraph{Motivation}
We address the construction of potentials for the gradient, the curl and the divergence operators, and more generally the exterior derivative, over triangulated domains. The potential operators are ought to satisfy bounds in Lebesgue norms that are explicitly computable in terms of the geometry. 
In particular, these bounds on the operator norms yield computable upper bounds for Poincar\'e--Friedrichs inequalities. 
Equivalently, this gives lower bounds for the eigenvalues of several vector Laplacians.

For example, if $\bff \in \bfL^{2}(\Omega)$ is a square-integrable vector field over some domain $\Omega \subseteq \bbR^3$, 
we want to find a potential $\bfu \in \bfH(\curl,\Omega)$ under the curl operator whose $\bfL^2$ norm is bounded in terms of the source field. This is expressed via the Poincar\'e--Friedrichs inequality 
\begin{align*}
    \min\limits_{ \substack{ \bfv \in \bfL^{2} \\ \curl \bfv = \bff } } \| \bfv \|_{\bfL^{2}(\Omega)}
    \leq 
    C_{\curl}
    \| \bff \|_{\bfL^{2}(\Omega)}
    .
\end{align*}
The constant in this inequality quantifies the stability properties of the Maxwell system on that domain. Any upper bound for it corresponds to a lower bound for the Maxwell eigenvalues.

Finding a potential operator, whether it be linear or nonlinear, and bounding the Poincar\'e--Friedrichs constant is generally not an easy undertaking. 
We address this topic in the special case where $\Omega$ admits a shellable triangulation.
Even though only contractible domains can ever admit a shellable triangulation, 
having computable upper bounds for such domains is an important step stone towards the more general case. 
For example, local patches around a simplex within a larger triangulation are shellable.
\\

@@

The remainder of this manuscript is structured as follows.
We review basic notions of triangulations in Section~\ref{section:triangulations} and provide computable upper bounds for Poincar\'e constants over triangulated domains in Section~\ref{section:poincare}.
We then review the notion of shellable triangulations of manifolds in Section~\ref{section:advancedtriangulations}.
We review the Sobolev spaces in vector calculus and the calculus of differential forms in Section~\ref{section:calculus}.
Bogovski\u{\i}-type integral operators over convex sets are discussed in Section~\ref{section:bogovskii}.
We construct an important geometric reflection operator in Section~\ref{section:extension}
and provide computable upper bounds for Poincar\'e--Friedrichs constants over shellable triangulations in Section~\ref{section:poincarefriedrichs}.

@@

The strictly positive convex combinations of the simplex' vertices are the \textit{interior} of the simplex,
and its remaining points constitute the \textit{boundary} of the simplex.
\\

@@

\begin{align}
    v_0 + \langle v_k - v_0, v_{k-1} - v_0, \dots, v_1 - v_0 \rangle,
    \\
    \langle v_n - v_{k+1}, \dots, v_n - v_{n-1} \rangle + v_{0},
    \\
    v_0 + \sum_{i=1}^{k} \lambda_{i} ( v_i - v_0 ),
    \\
    v_0 + \sum_{i=k+1}^{n} \lambda_{i} ( v_n - v_i ),
    \\
    v_0 + \sum_{i=1}^{k} \lambda_{i} ( v_i - v_0 ) + \sum_{i=k+1}^{n} \lambda_{i} ( v_n - v_i ),
\end{align}
\begin{align}
    v_0 \left( 1 - \sum_{i=1}^{k} \lambda_{i} \right)
    +
    \sum_{i=1}^{k} \lambda_{i} v_i 
    +
    \sum_{i=k+1}^{n-1} \lambda_{i} v_n
    -
    \sum_{i=k+1}^{n-1} \lambda_{i} v_i
\end{align}
\begin{align}
    v_0 \left( 1 - \sum_{i=1}^{k} \lambda_{i} \right)
    +
    \sum_{i=1}^{k} \lambda_{i} v_i 
    +
    \sum_{i=k+1}^{n-1} (-\lambda_{i}) v_i
    +
    \sum_{i=k+1}^{n-1} \lambda_{i} v_n
\end{align}

>>>>

\begin{abstract}
    We construct upper bounds for Poincar\'e--Friedrichs constants over triangulated domains. 
    This includes lower estimates for the singular values of the gradient, curl, and divergence operators as special cases. 
    The main application that we envision are computable estimates over finite element patches. 
\end{abstract}

@@

\paragraph{Motivation}
We address the construction of potentials for the gradient, the curl and the divergence operators, and more generally the exterior derivative, over triangulated domains. The potential operators satisfy bounds in Lebesgue norms that are explicitly computable in terms of the geometry. 
In particular, these bounds on the operator norms yield computable upper bounds for Poincar\'e--Friedrichs inequalities. 
Equivalently, this gives lower bounds for the eigenvalues of several vector Laplacians.

For example, if $\bff \in \bfL^{2}(\Omega)$ is a square-integrable vector field on some domain $\Omega \subseteq \bbR^3$, 
we want to find a potential $\bfu \in \bfH(\curl,\Omega)$ under the curl operator whose $\bfL^2$ norm is bounded in terms of the source field. This is expressed via the Poincar\'e--Friedrichs inequality 
\begin{align*}
    \min\limits_{ \substack{ \bfv \in \bfL^{2} \\ \curl \bfv = \bff } } \| \bfv \|_{\bfL^{2}(\Omega)}
    \leq 
    C_{\curl}
    \| \bff \|_{\bfL^{2}(\Omega)}
    .
\end{align*}
The constant in this inequality quantifies the stability properties of the Maxwell system on that domain. Any upper bound for it corresponds to a lower bound for the Maxwell eigenvalues.

Finding a potential operator, whether linear or non-linear, and bounding the Poincar\'e--Friedrichs constant is generally not an easy undertaking. 
We address this topic in the special case where $\Omega$ admits a shellable triangulation.
Even though only contractible domains can ever admit a shellable triangulation, 
having computable upper bounds for such domains is an important step stone towards the more general case. 
For example, local patches around a simplex within a larger triangulation are shellable.
\\

@@

The remainder of this manuscript is structured as follows.
We review basic notions of triangulations in Section~\ref{section:triangulations} and provide computable upper bounds for Poincar\'e constants over triangulated domains in Section~\ref{section:poincare}.
We then review the notion of shellable triangulations of manifolds in Section~\ref{section:advancedtriangulations}.
We review the Sobolev spaces in vector calculus and the calculus of differential forms in Section~\ref{section:calculus}.
Bogovski\u{i}-type integral operators over convex sets are discussed in Section~\ref{section:bogovskii}.
We construct an important geometric reflection operator in Section~\ref{section:extension}
and provide computable upper bounds for Poincar\'e--Friedrichs constants over shellable triangulations in Section~\ref{section:poincarefriedrichs}.

One component of our efforts are estimates for Bogovski\u{i}-type potential operators over convex domains, seen as operators between Lebesgue spaces. 
This follows closely the discussions in the literature~\cite{costabel2010bogovskiui} but no operator norm estimates are given in the literature, to the best of our best knowledge.
The Bogovski\u{i}-type potential operator first appeared for the mathematical discussion of hydrodynamical models~\cite{bogovskii1979solution} and has since been extended to exterior calculus. 
It can be identified as a formal adjoint to the Poincar\'e operator in exterior calculus. 
Our variant of the Bogovski\u{i} potential operator is simplified in comparison earlier discussions:
we only view it over convex domains and use simpler (that is, constant) weight function. 
The resulting operator has generally lower regularity properties in comparison to the regularized potential in the literature but the regularity is sufficient for our purposes and the operator norms are easily computed. We consider this outcome to be of independent interest.

The Bogovski\u{i} potential operator is repurposed to prove Poincar\'e--Friedrichs inequalities for differential forms over simplices and subject to boundary conditions. For example, given the curl of a vector field over a tetrahedron that vanishes along, say, two of the tetrahedron faces, we want to find the unique vector field potential. We achieve by constructing an auxiliary problem subject to full boundary conditions.

This construction of potentials requires triangulations with special conditions: 
we assemble the triangulation sequentially, adding one simplex after the other, such that the interface between new previous intermediate triangulation and the next simplex is always a boundary submanifold of either of them.  Loosely speaking, this guarantees that the interface with each successive simplex is sufficiently well-behaved. Such triangulations are also known as shellable. Shellability of simplicial complexes and, more generally, of polytopal complexes is a well-established notion in discrete geometry and combinatorics~\cite{ziegler2012lectures}, which we only employ in a simple special case. In particular, local stars in 2D and 3D are shellable.

Though we present our results in vector calculus, our main arguments are given in the formalism of exterior calculus. Exterior calculus has found widespread adoption in the theoretical and numerical analysis for vector-valued partial differential equations~\cite{hiptmair2002finite,arnold2006finite}. Loosely speaking, exterior calculus is the calculus of antisymmetric tensor fields. The formalism is independent of the dimension and highlights structures commmon to the gradient, curl, and divergence operators. It also enables us to leverage on the rich literature of Sobolev spaces of differential forms.

\cite{vohralik2005discrete}

@@

The strictly positive convex combinations of the vertices of the simplex constitute the \textit{ interior} of the simplex,
and its remaining points constitute the \textit{boundary} of the simplex.
\\

@@

Let $T$ be an $n$-dimensional simplex with some subsimplex $F$ We show that there exists a hyperplane of codimension one that intersects $T$ only at the $F$ and such that $T$ is only on one side of the hyperplane.

\begin{lemma}\label{lemma:oppositesubsimplex}
    Let $v_0, v_1, \dots, v_n \in \bbR^n$ be the vertices of a simplex $T$.
    Let $F, F'$ be the two subsimplices of $T$ with respective sets of vertices $v_0,\dots,v_k$ and $v_{k+1},\dots,v_n$.
    Then $T$ lies on one side of the affine subspace
    \[
        A := v_0 + \operatorname{lin}\{ v_1 - v_0, \dots, v_k - v_0 \} +  \operatorname{lin}\{ v_n - v_{k+1}, \dots, v_n - v_{n-1} \}.
    \]
    Moreover, $F = T \cap A$.
\end{lemma}
\begin{proof}
    Evidently, $A$ has dimension $n-1$ and $F \cap A \cap T$.
    Let us now suppose that $x \in A \cap T$.
    For some real coefficients $\lambda_1,\dots,\lambda_{n-1} \in \bbR$, it holds that 
    \begin{align}
        x 
        &
        = 
        v_0 + \sum_{i=1}^{k} \lambda_{i} ( v_i - v_0 ) + \sum_{i=k+1}^{n-1} \lambda_{i} ( v_n - v_i )
        \\&
        = 
        \left( 1 - \sum_{i=1}^{k} \lambda_{i} \right) v_0 
        + 
        \sum_{i=1  }^{k  } \lambda_{i} v_i 
        + 
        \left( \sum_{i=k+1}^{n-1} \lambda_{i} \right) v_n
        + 
        \sum_{i=k+1}^{n-1} (-\lambda_{i}) v_i 
        .
    \end{align}
    We know $x$ can be written as a unique convex combination of the vertices of $T$.
    Hence the $\lambda_{k+1}, \dots, \lambda_{n-1}$ are non-positive. 
    If $x \notin F$, then at least one of these coefficients must be negative. 
    But then their sum must be negative, which is a contradiction. 
    The desired statement follows.         
\end{proof}
